---
title: "Hw2 S135"
author: "Callin Switzer"
date: "October 6, 2014"
output: pdf_document
---

Homework #2     		October 6, 2014

1.  Select two points randomly in a rectangle of length=2 and width=1.
Compute summary statistics & histogram as in class.  In particular,
what is the (estimated) mean distance between the points?\
```{r}
nsim <- 10000
set.seed(12345)
# get points
points <- data.frame(x1 = runif(n = nsim, min = 0, max = 2), 
                     y1 = runif(n = nsim, min = 0, max = 1), 
                     x2 = runif(n = nsim, min = 0, max = 2), 
                     y2 = runif(n = nsim, min = 0, max = 1))
# calculate distances: sqrt((x2 - x1)^2 + (y2 - y1)^2)
dist <- with(points, sqrt((x2 - x1)^2 + (y2 - y1)^2))
hist(dist)
abline(v  = mean(dist), col = "red")
mean(dist)

```

2.  Picture two unit squares sharing a side (like adjacent cells on
a chessboard).  Select a point randomly from each.  What is the 
(estimated) mean distance between the points?

```{r}

points <- data.frame(x1 = runif(n = nsim, min = 0, max = 1), 
                     y1 = runif(n = nsim, min = 0, max = 1), 
                     x2 = runif(n = nsim, min = 1, max = 2), 
                     y2 = runif(n = nsim, min = 0, max = 1))

dist <- with(points, sqrt((x2 - x1)^2 + (y2 - y1)^2))
hist(dist)
abline(v  = mean(dist), col = "red")
mean(dist)

```

3.  Picture two unit squares sharing a corner and diagonal line 
(again, as on a chessboard).  Select a point randomly from each.  
What is the (estimated) mean distance between the points?

```{r}
nsim <- 10000

points <- data.frame(x1 = runif(n = nsim, min = 1, max = 2), 
                     y1 = runif(n = nsim, min = 1, max = 2), 
                     x2 = runif(n = nsim, min = 0, max = 1), 
                     y2 = runif(n = nsim, min = 0, max = 1))

dist <- with(points, sqrt((x2 - x1)^2 + (y2 - y1)^2))
hist(dist)
#plot(points$x2,points$y2, xlim = c(0,2), ylim = c(0,2))
#points(points$x1,points$y1, col = "red")
abline(v  = mean(dist), col = "red")
mean(dist)

```

4.  Select two points randomly in the unit square.  Draw a line
segment connecting the two points.  Now select two more points 
randomly in the unit square.  Draw a new line segment connecting 
the two new points.  Estimate the probability that the two line
segments cross (that is, have a point in common).

```{r}
# 2 points in unit square

linCross <- function(o) {
    # First two models
     df1 <- data.frame(x=runif(2), y=runif(2))
     m1 <- lm(y~x, df1) # slope
     df2 <- data.frame(x=runif(2), y=runif(2))
     m2 <- lm(y~x, df2) # slope
     
     # Plot them to show the intersection visually    
     #plot(df1,xlim = c(0,1), ylim = c(0,1))
     #segments(df1$x[1], df1$y[1],df1$x[2], df1$y[2])
     #segments(df2$x[1], df2$y[1],df2$x[2], df2$y[2])
     #points(df2)
     
     # Now calculate it!    
     a <- coef(m1)-coef(m2)
     cm <- rbind(coef(m1),coef(m2)) # Coefficient matrix
     foo <- c(-solve(cbind(cm[,2],-1)) %*% cm[,1])
     
     #points(foo[1], foo[2])
     
     ## Now, if we project the lines onto the x and y axis, 
     ## the point of intersection must be between them
     
     #check xs
     xx1 <- max(df1$x[1], df1$x[2]) >= foo[1] & foo[1] >= min(df1$x[1], df1$x[2])
     #check xs on second line
     xx2 <- max(df2$x[1], df2$x[2]) >= foo[1] & foo[1] >= min(df2$x[1], df2$x[2])
     # check ys
     xx3 <- max(df1$y[1], df1$y[2]) >= foo[2] & foo[2] >= min(df1$y[1], df1$y[2])
     #check ys on second line
     xx4 <- max(df2$y[1], df2$y[2]) >= foo[2] & foo[2] >= min(df2$y[1], df2$y[2])
     
     sum(xx1, xx2, xx3, xx4) == 4 
}

lc <- replicate(10000, linCross())
sum(lc) / 10000

```


5.  Select two points randomly in the unit disk (radius=1).
Compute summary statistics & histogram as in class.  In particular,
what is the (estimated) mean distance between the points?

```{r}
ddist <- function(o){
     nsim <- 2
     rdisk <- function(q)         # how to generate random points in the unit disk?
     {
       R <- matrix(runif(nsim), ncol=1)
       Theta <- matrix(runif(nsim,0,2*pi), ncol=1)
       XY <- cbind(R^q*cos(Theta),R^q*sin(Theta))
       #x11(width=5,height=5.5)    # keep old plot and open a new plotting window
       #plot(XY,xlim=c(-1,1),ylim=c(-1,1))
       #upp.crcl <- function(x) sqrt(1-x^2)
       #low.crcl <- function(x) -sqrt(1-x^2)
       #curve(upp.crcl,from=-1,to=1,add=T)
       #curve(low.crcl,from=-1,to=1,add=T)
       return(XY)
     }
     XY <- rdisk(1/2)
     return( with(points, sqrt((XY[2,1] - XY[1,1])^2 + (XY[2,2] - XY[1,2])^2)) )  
}

dists <- replicate(10000, ddist())
hist(dists)
mean(dists)

```

6.  Select three points randomly in the unit disk (radius=1).
Compute summary statistics & histogram as in class.  In particular, 
estimate the probability that the triangle is obtuse.

```{r}
nsim <- 3
rdisk <- function(q)         # how to generate random points in the unit disk?
{
  R <- matrix(runif(nsim), ncol=1)
  Theta <- matrix(runif(nsim,0,2*pi), ncol=1)
  XY <- cbind(R^q*cos(Theta),R^q*sin(Theta))
  #x11(width=5,height=5.5)    # keep old plot and open a new plotting window
  #plot(XY,xlim=c(-1,1),ylim=c(-1,1))
  #upp.crcl <- function(x) sqrt(1-x^2)
  #low.crcl <- function(x) -sqrt(1-x^2)
  #curve(upp.crcl,from=-1,to=1,add=T)
  #curve(low.crcl,from=-1,to=1,add=T)
  return(XY)
}
angl <- function(v,w) acos(v%*%w/sqrt((v%*%v)*(w%*%w)))
                 # dot (inner) product denoted by %*%

#XY <- rdisk(1/2) # here are the three points

#plot(XY, type = "l")

mxangl <- function(XY)
{
             #XY <- matrix(runif(6), ncol=2)
             A1 <- angl(XY[2,]-XY[1,],XY[3,]-XY[1,])
             A2 <- angl(XY[3,]-XY[2,],XY[1,]-XY[2,])
             A3 <- angl(XY[1,]-XY[3,],XY[2,]-XY[3,])
             max(A1,A2,A3) * (180/pi)

}

obt <- replicate(10000, mxangl(rdisk(1/2)) > 90)
sum(obt)/length(obt)
```

7. What is the cross-correlation between area & perimeter of random 
triangles in the unit disk?  Perform a linear regression as in class.
```{r}
# generate three points and calculate area vs. perimeter
side <- function(u) sqrt(u%*%u)   
angl <- function(v,w) acos(v%*%w/sqrt((v%*%v)*(w%*%w)))

rdiskAP <- function(q){
     nsim <- 3
     R <- matrix(runif(nsim), ncol=1)
     Theta <- matrix(runif(nsim,0,2*pi), ncol=1)
     XY <- cbind(R^q*cos(Theta),R^q*sin(Theta))
     S1 <- side(XY[3,]-XY[2,])
     S2 <- side(XY[1,]-XY[3,])
     S3 <- side(XY[2,]-XY[1,])
     A1 <- angl(XY[2,]-XY[1,],XY[3,]-XY[1,])
     area <- (1/2)*S2*S3*sin(A1)
     peri <- S1+S2+S3
     c(area=area,peri=peri)
}

#calculate area
AP <- t(replicate(10000, rdiskAP(1/2)))

plot(AP[,1], AP[,2])
modAP <- lm(AP[,2]~AP[,1])
abline(modAP, col = "red")
sm <- summary(modAP)
sqrt(sm$r.squared) # here is the correlation

```


8.  Select four points randomly in a disk.  With probability 1,
no three of the points are collinear, so the convex hull of the four
points is either a triangle (one point "inside" the others) or a 
quadrilateral.  Estimate the probability that the convex hull is 
a triangle.

```{r}
cnvx <- function(n)           # convex hull of n points in the unit disk (radius=1)
{
  R <- matrix(runif(n), ncol=1)
  Theta <- matrix(runif(n,0,2*pi), ncol=1)
  XY <- cbind(sqrt(R)*cos(Theta),sqrt(R)*sin(Theta))
  #x11(width=5,height=5.5)    # keep old plot and open a new plotting window
  #plot(XY,xlim=c(-1,1),ylim=c(-1,1),pch=22)
  h <- chull(XY)
  h <- c(h, h[1])
  #lines(XY[h, ],col='red',lwd=3)
  #upp.crcl <- function(x) sqrt(1-x^2)
  #low.crcl <- function(x) -sqrt(1-x^2)
  #curve(upp.crcl,from=-1,to=1,add=T)
  #curve(low.crcl,from=-1,to=1,add=T)
  length(unique(h)) == 4
}

chp <- replicate(10000, cnvx(4))
1 - sum(chp)/length(chp) # probability that it's a triangle

```

9.  Do the same as in the previous problem, except select the four
points randomly in an isosceles right triangular region.

```{r}
# select points from right triangular region

xyIs <- function(o){
     x <- runif(1)
     y <- runif(n = 1, min = 0, max = 1)
     #plot(x,y, xlim = c(0,1), ylim = c(0,1))
     #segments(0,1,1,0)
     if(y > -x + 1) {
          x <- 1-x 
          y <- 1-y
          }
     #points(x,y, pch = 4, col = "red")
     return(c(x=x, y = y))
}

prob4 <- function(o){
     xx <- t(replicate(4, xyIs()))
     h <- chull(xx)
     h <- c(h, h[1])
     #plot(xx[,1], xx[,2], xlim = c(0,1), ylim=c(0,1)); segments(0,1,1,0)
     #lines(xx[h, ],col='red',lwd=3)
     length(unique(h)) == 4
}


p4 <- replicate(10000, prob4())
1-sum(p4)/length(p4)
```

10. Ask an original question about geometric probability, and
answer it (approximately) using R's simulation capabilities.

```{r}
## what is the volume of a random sphere with radius~Unif(1)
volSph <- function(o){
     rad <- runif(1)
     vol <-(4/3) * pi * rad^3   
}

vs <- replicate(10000, volSph())
mean(vs)


```