Homework #3     		October 31, 2014

1.  Consider a standard deck of 52 playing cards and draw 5 cards at
random.  Write simulations that calculate the probability of:
     
a.  one pair (contains two cards of the same rank, plus three other 
                   unmatched cards)
```{r}
deck <- data.frame(num = rep(1:13, 4), 
                   suit = rep(c("heart", "spade", "club", "diamond"), each = 13))
Pair <- function(o){
     h <- sample(deck$num, size = 5, replace = FALSE)
     sum(table(h) == 2) == 1 # returns T if you have one pair
}

p <- mean(replicate(10000, Pair()))
p

```
                   
b.  two pairs (contains two cards of the same rank, plus two cards of 
               another rank that match each other but not the first pair, plus 
               one unmatched card).
               
```{r}
Pair2 <- function(o){
     h <- sample(deck$num, size = 5, replace = F)
     sum(table(h) == 2) == 2 # returns T if you have two pair
}

p2 <- mean(replicate(10000, Pair2()))
p2

```
c.  triple (contains three cards of the same rank, plus two other 
            unmatched cards)
            
```{r}
triple <- function(o){
     h <- sample(deck$num, size = 5, replace = F)
     sum(table(h) == 3) == 1 # returns T if you have a triple
}

t <- mean(replicate(10000, triple()))
t

```


2.  Devise simulations for your own favorite gambling games involving 
coins, dice, cards, shotguns, ...

```{r}
# flush
flush <- function(o){
     f <- sample(deck$suit, 5, replace = F)
     table(f)
     (sum(table(f) == 5) == 1) 
}

fl <- mean(replicate(10000, flush()))
fl

```
3.  Given two random positive integers less than N, what is the
probability that no factor exceeding 1 divides both of them? What
is the limit of this probability as N approaches infinity?
```{r}
# found this function on stack exchange
factors <- function(x) {
    x <- as.integer(x)
    div <- seq_len(abs(x)) # gives 1:x
    factors <- div[x %% div == 0] # remainder = 0
    factors <- list(factors)
    return(factors)
}

getInts <- function(N=100, o){
     n <- 1:N
     nums <- sample(N, 2, replace = F)
     # find all factors of N
     fac <- sapply(nums, factors)
     
     fac <- unlist(fac) # unlist
     fac <- fac[fac !=1] # remove 1
     #sum(table(fac) > 1) > 0 # returns true if there are duplicate factors
     sum(duplicated(fac)) == 0 # returns true if there are no duplicate factors
}



N <- 100000
GI <- function(N){mean(replicate(100000, getInts(N)))}
#probs <- sapply(N, GI)
#probs
# 0.60624


```

4.  Suppose a needle of length 1 is thrown at random on a plane 
marked by parallel lines of distance 1 apart.  What is the 
probability that the needle will intersect one of the lines?

```{r}
plotNeedle <- function(x = 1/2, y=1/2, r = 1) {
     x <- runif(1)
     y <- runif(1)
     angles <- seq(0,2*pi,length.out=360)
     angles <- sample(angles, 1)
     x2 <- r*cos(angles)+x
     y2 <- r*sin(angles)+y
     
     # I got rid of all the plotting stuff for the simulations
     #plot(x,y, xlim = c(-1,2), ylim = c(-1,2))
     #lines(x = c(x, x2),y = c(y,y2))
     #points(r*cos(angles)+x,r*sin(angles)+y)
     #abline(h = 1, lty = 5); abline(h = 0, lty = 5)
     # will be true if the needle crosses the line
     return(I(r*sin(angles)+y) > 1 | I(r*sin(angles)+y) < 0) 
}

#plot(runif(1), runif(1), type = "n",xlim = c(-1,2), ylim = c(-1,2) )
#abline(h = 1, lty = 5); abline(h = 0, lty = 5)
crosses <- replicate(100000, plotNeedle())
sum(crosses)/length(crosses)

# theoretical?
(2 + pi)/(3+2+pi) # = 0.6315 (not sure if this is right)

# answer from simulation: 0.635


```

5.  Browse John Maindonald's "Using R for Data Analysis and Graphics - 
Introduction, Examples and Commentary", which is available at:

http://wwwmaths.anu.edu.au/~johnm/r/

(Download and save the PDF file 'usingR.pdf'.  Datasets are found
in the '/dsets' subdirectory.)  I considered his book with John Braun
as a textbook for this course, but it did not discuss R programming
in enough depth.  Search for interesting data analysis exercises
and work them out!



I chose to do the following problem from 6.5: 
1. Using the data set painters (MASS package), apply principal components analysis to the scores for Composition, Drawing, Colour, and Expression. Examine the loadings on the first three principal components. Plot a scatterplot matrix of the first three principal components, using different colours or symbols to identify the different schools.

```{r, fig.height=5, fig.width=5}
library(MASS)
data(painters)
#?painters

paint1 <- scale(painters[,1:4])
paintPCA <- princomp(paint1)
#plot(paintPCA$scores[,1], paintPCA$scores[,2], col = painters$School)

#pairs(paintPCA$scores, col = painters$School)
pairs(paint1[,1:3], col = painters$School, pch = 20)

biplot(princomp(paint1[,1:4]), cex = 0.5, 
       col = c(rgb(0,0,0, 0.5), "red"), xlim = c(-0.4, 0.4))
```


I also chose this exercise from CH.8: 
14. Write an R function that simulates a student guessing at a multiple choice test consisting of 40 questions, where there is chance of 1 in 5 of getting the right answer to each question. Find the mean and variance of the student's answers. Compare with the theoretical values of .2 and .16.
```{r}
correctAnswer <- function(o) sample(1:5, size = 1)
guess <- function(o) sample(1:5, size = 1)

answers <- function(o){
     return(c(correctAnswer(), guess()))
}

ff <- function(o){
     corAndGuess <- replicate(40, answers())
     dim(corAndGuess) <- c(40, 2)
     su<-sum(corAndGuess[,1] == corAndGuess[,2])/40
     var<-var(corAndGuess[,1] == corAndGuess[,2])
     return(c(sum = su, var = var))
}

#mean(replicate(1000, ff())[1,]) # very close to 0.2
#mean(replicate(1000, ff())[2,]) # very close to 0.16
```


6.  Browse Venables & Smith's "An Introduction to R", which is
available at:
     
     http://cran.r-project.org/manuals.html

(Download and save the PDF file 'R-intro.pdf'.)  This is more
abstract and has no exercises.  I would have liked to find a book 
halfway between Maindonald (which is easy) and Venables & Smith
(which is hard).

7.  Likewise, browse Green's "The R Primer"

http://www.stat.washington.edu/cggreen/rprimer/rprimer.html

and look for topics that interest you.  Again, there are no
exercises.

8.  In class, we examined M/M/1 and M/M/2 in files r7.R and r8.R.  
Remember that the "system" contains both the queue (waiting room) 
and service (doctor's office).  Previously we computed statistics 
and drew plots associated with:  
     Wq = time in queue per patient, 
     Ls = number of patients in system. 
Do the same for:   
     Ws = time in system per patient, 
     Lq = number of patients in queue
in the case of one doctor and the case of two doctors.  
(Just start with files r7.R & r8.R and edit appropriately.)
                                                         
                                                         
```{r}
N <- 24*7      # number of hours in a week
C <- 1         # number of doctors    
A <- 2         # mean rate of patient arrivals per hour
B <- 3         # mean rate of patient treatments per hour,
               #   per busy doctor
M <- 250       # number of iterations (weeks)
# Wq           # waiting time in queue per patient
               #   (excludes service time)
# Ls           # number of patients in system (both queue and service)

# Ws = Wq + treatment time
# Lq = Number of patiens in system - number getting treated

emrgncy.week <- function(o)
  {
    K <- rpois(1,N*A)      # number of patients in week o
    P <- matrix(0,K,3) 
    P[,1] <- sort(runif(K,0,N)) # gives times of arrival
    P[,3] <- rexp(K,B)     # treatment length is Exp[B]
    D <- rep(0,C)          # initially all 0s (Is the doctor available?)
    for (i in 1:K)         # begin small loop
      {
        j <- which.min(D) # which doctor is next available
        P[i,2] <- max(P[i,1],D[j]) #P[,2] is the "treatment begin time"
        D[j] <- P[i,2] + P[i,3]
      }                    # end small loop
    P
}                          # end definition of emrgncy.week

#emrgncy.week()

waiting.list <- function(o)
  {
    P <- emrgncy.week()
    Wq <- P[,2]-P[,1]    # P[,2]-P[,1] is of indeterminate length K   
    Ls <- sapply(1:N, function(t) length(P[P[,1]<t & t<P[,2]+P[,3],1]))
    list(Wq + P[,3],ifelse(Ls < 2, 0, Ls - 2))
  }

#waiting.list()

X <- sapply(1:M, function(o) waiting.list())
#X
Ws <- unlist(X[1,])
Lq <- unlist(X[2,])


# I need to check these
mean(Ws)
mean(Lq)



```

9.  A new museum opens, but tickets must be purchased in advance.
People buy tickets according to a Poisson process at rate 500 (per day). 
For each buyer, the separation S between their purchase time
and their museum visit is Unif[0,4] (in days).  What is the expected
cumulative number of visitors T days after opening, where T=1, 2, 3, 
4, 5, 6 or 7?

```{r}
# this one is probably wrong...

rate <- 500

K <- rpois(7,rate)      # number of tickets bought on day o
visits <- numeric()                             # when each of those people visit the museum
i <- 1:length(K)
ss <- function(i) {
     K <- rpois(7, rate)
     sum(K[1:i] * 1/4)   
}
ss(1)
ssV <- Vectorize(ss)


stuff <- t(replicate(1000, ssV(i)))

plot(colMeans(stuff), xlab = "day", ylab = "visitors") # this shows cumulative visitors
cumsum(colMeans(stuff))/2 # not right

```

10.  A company has a large number of machines that tend to break
down according to a Poisson process at rate 5 (per hour).  A single
repair person is available.  Due to the nature of the machines,
it takes the person either 9 minutes or 12 minutes to repair a
machine.  The 12-minute type of malfunction occurs with probability
1/3.  What is the expected number of machines down at any given
time?  Also, what is the probability that more than three machines
will be down at any given time?

```{r}
rate <- 5  # 5 breakdowns per hour  == 5/60 breakdowns per minute
machBreak <- function(o){
     breaks <- rpois(60, 5/60) # number of breaks per minute for an hour
     timeFixing <- numeric()
     machineBroken <- numeric()
     timeFix <- numeric()
     for(i in 1:length(breaks)){
          if(timeFix[1] == 0 & !is.na(timeFix[1])) timeFix <- timeFix[-1]
          #timeFix
          #timeFixing[i] 
          longShort <- rbinom(n = breaks[i], size = 1, prob = (1/3))
          #longShort
          fixThisRound <- ifelse(longShort == 0, 9, 12)
          timeFix <- c(timeFix, fixThisRound)
          if(!is.na(timeFix[1])) timeFix[1] <- timeFix[1] - 1
          #timeFix
          machineBroken[i] <- ifelse(is.na(timeFix[1]), 0 , length(timeFix[timeFix !=0]))
          #totalFixTime <- sum(timeFix)
          #print(c(time = timeFix, machinesBroken = machineBroken[i]))
          #i = i + 1
     }
     return(mean(machineBroken))
}

#meanBrokenMachines <- mean(replicate(10000, machBreak()))
#meanBrokenMachines # 1.091588


# Probability that > 3 machines will be broken down
machBreakProb <- function(o){
     breaks <- rpois(60, 5/60) # number of breaks per minute for an hour
     timeFixing <- numeric()
     machineBroken <- numeric()
     timeFix <- numeric()
     for(i in 1:length(breaks)){
          if(timeFix[1] == 0 & !is.na(timeFix[1])) timeFix <- timeFix[-1]
          #timeFix
          #timeFixing[i] 
          longShort <- rbinom(n = breaks[i], size = 1, prob = (1/3))
          #longShort
          fixThisRound <- ifelse(longShort == 0, 9, 12)
          timeFix <- c(timeFix, fixThisRound)
          if(!is.na(timeFix[1])) timeFix[1] <- timeFix[1] - 1
          #timeFix
          machineBroken[i] <- ifelse(is.na(timeFix[1]), 0 , length(timeFix[timeFix !=0]))
          #totalFixTime <- sum(timeFix)
          #print(c(time = timeFix, machinesBroken = machineBroken[i]))
          #i = i + 1
     }
     return(sum(machineBroken > 3)/length(machineBroken))
}

#Prob <- mean(replicate(10000, machBreakProb()))
#Prob # 0.046 is the prob of > three broken machines at any given time

```

THE FOLLOWING FOUR PROBLEMS ARE HARD!
     
     11.  Consider the one-dimensional interval [0,x] with x > 1. 
Imagine it to be a street for which parking is permitted on one
side.  Cars of unit length are one-by-one parked completely at
random on the street and obviously no overlap is allowed with
cars already in place.  What is the mean number, M(x), of 
cars that can fit?  What is the limit of M(x)/x as x approaches
infinity?



```{r}
X <- 10 # length of parking interval
plot(1,1,type = "n", xlim = c(0, X), ylim = c(0,0.3))

# park first car
bump <- function(o) runif(1,0, X-1)
x1 <- bump()
car <- c(x1, x1 + 1)
car <- function(o){
     x1 <- bump()
     c(x1, x1 + 1)
}

car()

lines(x = car(), y = c(0.02, 0.02) + runif(1,min = 0.01, 0.05), col = sample(1:3, 1), lwd = 5)

nsim <- 100
prk <- matrix(data = 0,nrow = nsim, ncol = 2)
X <- 5 # length of parking interval
plot(1,1,type = "n", xlim = c(0, X), ylim = c(0,0.3))
i = 1
for(i in 1:nsim){
     
     newCar <- car()
     if(sum(newCar[1] > prk[,1] & newCar[1] < prk[,2]) == 0 &
             sum(newCar[2] > prk[,1] & newCar[2] < prk[,2]) == 0){
         prk[i,] <- newCar
         lines(x = newCar, 
               y = c(0.02, 0.02) + runif(1,min = 0.01, 0.05), 
               col = sample(1:3, 1), lwd = 1)
     }
     prk
     i = i + 1
}
# find space between each of the cars
prkO <- prk[prk[,2] != 0, ]
prkO <- prkO[order(prkO[,1],decreasing = F), ]
dfDist <- data.frame(c(0, prkO[,1]+1), c(prkO[,1], X))
max(dfDist[2]- dfDist[1]) > 1 # only keep the ones where this is FALSE


# idea
vec <- sample(x = c(0,1), size = 10000, replace = T, prob = c(0.6, 0.4))
k <- 1:(10000 -2)

fii <- vec[k]*vec[k+1]*vec[k+2]
sum(fii == 1)/length(vec)

# park another car



```

12.  Consider a variation of the above, in which the interval [0,x]
is replaced by the discrete linear lattice 1, 2, 3, ..., n.
Each car is a line segment of unit length and covers two lattice
points when it parks.  No car is permitted to touch points that
have already been covered.  The random parking process stops when 
no two adjacent pairs of lattice points are left uncovered.  What 
is the analogous limiting density as n approaches infinity?

13. MORE COINS

Suppose you want to play a game that has the following rules:
     
     A. Each play of the game involves repeatedly flipping three unbiased 
independent coins. You keep track, for each coin k, the difference d.k
between the number of heads tossed and the number of tails tossed.
The game stops when d.1 = 0, or d.2 = 0, or d.3 = 0.

B. If you decide to play the game, you are required to pay $1 for each
flip of the three coins. You are not allowed to quit during a play of 
the game.

C. You receive $4 at the end of each play of the game.

Thus you win money if the number of flips required is fewer than 4, but you
lose money if more than 4 flips are required. What is the expected payoff
(or loss) when playing this game?


14. MORE DICE

Background: opposite sides of a cubic die always sum to 7,
that is, 1 is opposite 6, 2 is opposite 5, and 3 is opposite 4.

Toss an unbiased die repeatedly, keeping track of three things:
     
     A. the difference d.1 between the number of 1s tossed and
the number of 6s tossed;

B. the difference d.2 between the number of 2s tossed and
the number of 5s tossed;

C. the difference d.3 between the number of 3s tossed and
the number of 4s tossed.

What is the probability that, at some point in time, we have
d.1 = d.2 = d.3 = 0 simultaneously?  What is the median
duration of the game?

